---
title: "DRAFT OF: Cookie Clicker Bot"
author: "Mac Strelioff"
date: "`r Sys.time()`"
output:
  blogdown::html_page:
    toc: true
math: true
---

```{r}
library(reticulate)
```

```{r,include=FALSE}
knitr::opts_chunk$set(comment = ">>>")
#py_discover_config() # to see versions of Python
use_python("/anaconda3/bin/python")
# py_config() # check paths
# test imports
#import("numpy",as="np")
#import("requests")

# path to interpreter from python (I ran in a juperter nb to find the path used here)
# import sys
# print(sys.executable)

# for figures, without issues..;
# https://blogdown-demo.rbind.io/2017/10/06/adding-figures-and-other-images-to-posts/
```


# Overview






# Background and Setup 

- I was interested in developing programs that can intelligently interact with a web interface.. 
- I'll highlight some bits of code, but a Jupyter notebook with all code to run these algorithms can be found (HERE) -- link to my git repo with the notebook...

## Cookie Clicker Game

- pic of the interface
- You get cookies for clicking the big cookie. You can spend these cookies on buildings, which generate a number of cookies per second. Buildings get more expensive when bought, so the marginal benefit of any particular building depends on how many are currently owned. 

# Strategies

## Naive

Simple to implement, but probably not the best.

## Max revenue per second

## Min wait time until the highest rps option

- rationale: differences in revenue per building are quite large, so ... .
- pics from ipad of concept figure
- proof that balance doesn't matter 
- sketches of preference as rps and cost of another building change. 

# Performance 

- revenue per second across clicks (proxy for time)

- building counts per across clicks (proxy for time)




