---
title: 
draft: yes
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
menu:
  data-science:
    parent: Learn Data Science
    weight: 1
linktitle: Working Environment
#toc: yes
type: docs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE,
                      cache=TRUE,
                      comment = '')

library(osfr) # for interfacing with OSF projects through R
library(lme4) # for common frequentist models
library(brms) # for common bayesian models
library(png)

# set up to work with python
#setwd("/anaconda3") # need to set wd to /anaconda3 for access to packages installed under this distribution
#library(reticulate) # for interfacing with python
# use_condaenv(condaenv = "r-nlp", conda = "/anaconda3/bin/conda")
# use_python("/anaconda3/bin/python")#,required = TRUE)
# py_discover_config()
# ** for some reason these break compilation in papaja
# py_config() # check that anaconda distribution is used
# import("pandas")

# was in its own r code chunk
# can also use 
# # py_run_file()
# py_run_string("
# import pandas as pd
# c={'col1': [1, 2], 'col2': [3, 4]}
# c = pd.DataFrame(data=c)
# ")
# use py$<var> to access vars from python
# py$c

set.seed(10201991) # set seed for reproducability
```

<!---
<iframe width="560" height="315" src="https://www.youtube.com/embed/LgYl1ffS_6Y" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
--->

# Working Environment

My working environment relies on bash and git. Here I provide some background on these tools and a walkthrough for automatic, scheduled pushes to git.

## Working With Bash and Git

### Useful Bash Commands

```{bash, eval=FALSE}
login     # prompts for username and then password to log in
man       # man (command) returns documentation a command
```

### Directory Commands

```{bash, eval=FALSE}
pwd       # print working directory
cd        # changes directory to home
cd \      # changes directory to (/path)
cd ..     # moves up one folder in the directory
ls -al    # lists contencts of directory, -a includes hidden items, -l includes details
```

### File Operations

```{bash,eval=FALSE}
mkdir    # make a new directory
touch    # create a file
open     # open a file
mv       # move (file) (to), can also be used to rename a file
cp       # copy (file) (to) 
rm       # delete a file
```

### Other Utilities
```{bash,eval=FALSE}
echo    # print commands screen
date    # print date
cal     # print calendar (year)
```


## Creating Custom Commands

This section is partially based on [this Medium blog post](https://medium.com/devnetwork/how-to-create-your-own-custom-terminal-commands-c5008782a78e).

Use the alias command to map an execution of bash commands to a single command. For example, this will create a command 'hi' where the Luca voice slowly says 'hello world';
```{bash, eval=FALSE}
alias hi="say -v Luca  -r 100 'hello world'"
hi # test this alias
```

Functions can be created with the following syntax; 
```{bash, eval=FALSE}
function function_name() {
 <args> 
}
```

Inputs are denoted with \$1, \$2, ... . For example, when called from the terminal, this function will print 'Your Input Was: ' followed by the first argument it receives;
```{bash, eval=FALSE}
function print_my_input() {
 echo 'Your Input Was:' $1 
}
```

To create a file where custom commands are defined, navigate to the home directory, create, and open the file that will hold function definitions;
```{bash, eval=FALSE}
cd 
touch .custom_commands
open .custom_commands
```

Define commands you wish to use in future Terminal sessions in the .custom_commands file. The next section describes how to source this file automatically when a Terminal session is initialized. 

## Sourcing Files During Terminal Initialization
This section is partially based on [this StackOverflow post](https://stackoverflow.com/questions/19662713/where-do-i-find-the-bashrc-file-on-mac)

Bash sources a run control file (~/.bashrc) at initialization. Check if this file exists in the home directory with;
```{bash, eval=FALSE}
cd
ls -a
```

If there is no .bashrc file listed, then first open the .bash_profile file in a text editor and add the following;
```{bash, eval=FALSE}
if [ -f ~/.bashrc ]; then
    . ~/.bashrc
fi
```

Then create and open a .bashrc file in the home directory with;
```{bash, eval=FALSE}
cd
touch .bashrc
open .bashrc
```

In a text editor add the following lines to the .bashrc file;
```{bash, eval=FALSE}
# Source global definitions
if [ -f /etc/bashrc ]; then
    . /etc/bashrc
fi

# source file with function definitions
source ~/.custom_commands
```

This last line will source the .custom_commands file described at the end of the previous section, thereby defining any aliases or functions in the .custom_commands file whenever a Terminal session is initialized.

## Scheduling Bash Functions With Crontab

Some basic crontab commands; 
```{bash,eval=FALSE}
crontab -l # lists crontab jobs
crontab -r # removes all crontab jobs
crontab -e # edits list of crontab jobs
```

The crontab -e command is somewhat hard to navigate, the following code is an easier way to create crontab jobs from ther terminal; 
```{bash, eval=FALSE}
crontab -l | { cat; echo "* * * * * some entry"; } | crontab -

# format for string argument for crontab;
* * * * * "some entry"
- - - - -
| | | | |
| | | | ----- Day of week (0 - 7) (Sunday=0 or 7)
| | | ------- Month (1 - 12)
| | --------- Day of month (1 - 31)
| ----------- Hour (0 - 23)
------------- Minute (0 - 59)
```

I simplify this farther in a custom function called 'schedule' that I define in \ref{sec: schedule}.

To execute functions from a user, the user must be specified in /etc/cron.allow -- however, this file can only be edited by a superuser. To create and edit this file, use; 
```{bash, eval=FALSE}
cd /etc
sudo touch cron.allow
sudo nano cron.allow
```

nano opens a text editor in ther terminal, and that can be used to edit and save changes to cron.allow. 

More information on crontab [here](https://www.computerhope.com/unix/ucrontab.htm).

## Git and GitHub

### Setup
```{bash,eval=FALSE}
# create user name and email for tagging contributions later
git config --global user.name "USERNAME"   
git config --global user.email "EMAIL" 
```

Connecting to Git can be done through a username/password or through SSH verification. 

### Username/Password

This didn't work for ubdates scheduled through crontab. 

### SSH

First, [check for SSH keys with](https://help.github.com/articles/checking-for-existing-ssh-keys/);
```{bash,eval=FALSE}
cd ~/.ssh  # navigate to user/.ssh folder
ls -a      # list all files
```

A public SSH key is in id_rsa.pub, which can be opened in a text editor. 

If no id_rsa file exists, or if you wish to create a new SSH key, you can use the following command to generate a new SSH key;
```{bash,eval=FALSE}
ssh-keygen -t rsa -b 4096 -C "your_email@example.com" # run this if id_rsa file does not exist
```
This will prompt you for the file location where the SSH key should be saved -- id_rsa is the default file. 

From here, follow the steps [to add an SSH key](https://help.github.com/articles/generating-a-new-ssh-key-and-adding-it-to-the-ssh-agent/), summarized below. 

First, create a config file for the SSH agent with; 
```{bash,eval=FALSE}
touch ~/.ssh/config
```

Open the config file in a text editor and add the the following commands; 
```{bash,eval=FALSE}
Host *
 AddKeysToAgent yes
 UseKeychain yes
 IdentityFile ~/.ssh/id_rsa
```

Next, from Terminal, add the key from the file (e.g. id_rsa) to the ssh agent with; 
```{bash,eval=FALSE}
ssh-add -K ~/.ssh/id_rsa
```

Finally, [add the ssh key to the GitHub account](https://help.github.com/articles/adding-a-new-ssh-key-to-your-github-account/).

And [test the connection](https://help.github.com/articles/testing-your-ssh-connection/) with; 
```{bash,eval=FALSE}
ssh -T git@github.com
```

### Working From a Repository

Create a repository through GitHub's user interface (log in and select 'new repository'). 

Then create a directory on a local computer, possibly using bash commands; mkdir and cd.

The simplest way is to clone the repo from GitHub with;
```{bash,eval=FALSE}
git clone (url)     # this will clone from the url to the current directory
```

Another way, from inside the directory that is to be the local repo, issue these commands;
```{bash,eval=FALSE}
git init   # initialize a repo
git remote add origin (url or SSH address) # point local repo to remote repo
```

### Switching between SSH and HTTPS Access

In both examples below, replace 'USERNAME/REPOSITORY' with the appropriate values. From within the local repo folder,

Change from HTTPS to SSH with; 
```{bash,eval=FALSE}
git remote set-url origin git@github.com:USERNAME/REPOSITORY.git
git remote -v # list fetch and push destinations, should start with git@github.com
```

Change from SSH to HTTPS with;
```{bash,eval=FALSE}
git remote set-url origin https://github.com/USERNAME/REPOSITORY.git
git remote -v # list fetch and push destinations, should start with https://github.com/
```

[More detailed guide here](https://help.github.com/articles/changing-a-remote-s-url/).

Other methods for addressing 'Device not configured' errors [here](https://stackoverflow.com/questions/40274484/fatal-could-not-read-username-for-https-github-com-device-not-configured). 

### Repository Maintanence And Collaboration

```{r}
# download.file("https://i.stack.imgur.com/nWYnQ.png",'y.jpg', mode = 'wb')
jj <- readPNG("y.jpg",native=TRUE)
plot(0:1,0:1,type="n",ann=FALSE,axes=FALSE)
rasterImage(jj,0,0,1,1)
```

Useful commands;
```{bash,eval=FALSE}
git pull    # adds files from the configured remote repo to your local repo and workspace
git add .   # adds all new files to be tracked
git add -u  # updates all files
git add -A  # does both '.' and '-u'
git commit -m "message" # saves changes local repo
git push    # updates repo on GitHub
```

Collaborators can all access the GitHub repo through push/pull commands. In collaborative settings you might need to pull, so that your repository is current, before pushing changes from the local repository to the remote repository on GitHub. 

## Useful Custom Aliases and Functions

#### gitup: combines add, commit, and push into one command

```{bash,eval=FALSE}
alias gitup="git add -A; git commit -m 'auto'; git push"
```

#### gitupall: updates all git repos in my git folder

```{bash,eval=FALSE}
function gitupall{
for filename in /Users/mac/git/*; do
echo "**************************updating $filename"
cd "$filename"; 
gitup
cd ..
done
}
```

#### schedule: schedules a command to be executed periodically

```{bash, eval=FALSE}
function schedule(){
crontab -l | { cat; echo "$1"; } | crontab -
}

# format for string argument for crontab;
* * * * * "command to be executed"
- - - - -
| | | | |
| | | | ----- Day of week (0 - 7) (Sunday=0 or 7)
| | | ------- Month (1 - 12)
| | --------- Day of month (1 - 31)
| ----------- Hour (0 - 23)
------------- Minute (0 - 59)
```

Example use;
```{bash, eval=FALSE}
# e.g. use: schedule updating of all repos each hour
schedule "0 0-23 * * * gitupall"
crontab -l # lists crontab jobs after schedule call
```


## My Setup

### Bash Aliases, Functions, Scheduled Commands

In my root directory, I have a file called .custom_commands with contents; 
```{bash,eval=FALSE}
# aliases

# goes to my folder of git projects
alias gitgo="cd /Users/mac/git; ls"
echo 'sourced alias gitgo'

# pushes updates on a git project
alias gitup="git add -A; git commit -m 'auto'; git push"
echo 'sourced alias gitup'

# functions

# schedule "(m h DoM M DoW command)"
# schedules command to be executed at 
# minute m of hour h on DoM days of the month in months M or DoW days of the week.
function schedule(){
crontab -l | { cat; echo "$1"; } | crontab -
}
echo 'sourced function schedule'

# for updating all git projects, schedule this to be done periodically
function gitupall(){
for filename in /Users/mac/git/*; do
echo "************************** updating $filename"
cd "$filename";
git remote -v    # prints the fetch and push destinations
gitup
done
cd # return to home directory
}
echo 'sourced function gitupall'
```

I also have a .bashrc file in my root directory, which sources the custom commands. Particularly the contents of the .bashrc file are;

```{bash,eval=FALSE}
# Source system-wide definitions
if [ -f /etc/bashrc ]; then
    . /etc/bashrc
fi

# source file with function definitions
source ~/.custom_commands
```

I also have a file in my root directory for scheduling, called .scheduled with contents; 
```{bash, eval=FALSE}
# put all scheduled functions here

# source custom commands
source ~/.custom_commands

# activate key agent
eval "$(ssh-agent -s)"

# test SSH key / print username
ssh -T git@github.com

# update all git repos
gitupall
```

I then scheduled this file to be read hourly by opening a terminal session, which sources .bashrc at initialization thereby sourcing my custom commands, and running; 
```{bash,eval=FALSE}
schedule "0 0-23 * * * source ~/.scheduled"
crontab -l # the scheduled job should be listed as a crontab job
```

